---
title: "P8105 Data Wrangling 2"
author: 'Ying Chen (UNI: yc384)'
date: "9/19/2019"
output: github_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
options(tibble.print_min = 3)
```

## Import datasets

```{r manipulation}

## Codes run in class

litters_data = read_csv("./data/FAS_litters.csv", col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv", col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```

## piping: %>%
## came out in 2014
## one line code does not recommend - hard to follow
## step by step codes are easier to follow and read
## table in and table out of tidyverse works well with piping ...

## Selecting
```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)

## select all vars starts with pups
select(litters_data, litter_number, gd0_weight, starts_with("pups"))

## starts with litter_number, group and everthing else
select(litters_data, litter_number, group, everything())
subdata=select(litters_data, litter_number, group)
select(litters_data, -group)
select(litters_data, group:pups_born_alive)
select(litters_data, -pups_survive)

```

## rename varaibles: not using very often
## only rename it when the name is too long
```{r}
rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)

```

## filter
## == is separate operator than = (assign)
```{r}
filter(litters_data, group == "Con7")
filter(litters_data, group == "Mod8")

filter(litters_data, gd_of_birth == 20)
filter(litters_data, pups_born_alive < 6, group == "Con7")
filter(litters_data, pups_born_alive < 6 | group == "Con7")

filter(litters_data, group %in% c("Con7", "Mod8"))

filter(litters_data, pups_born_alive >= 4, pups_born_alive <=6)
filter(litters_data, pups_born_alive * 2 >= 6)

## use this type of deleting missing cells
drop_na(litters_data, gd0_weight)

```

## !(pups_survive == 4) ! can be put outside the (), works same as !=

## Mutating is great!
## mutate: creating and modifying new varaibles
```{r}
mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight, #create new
  group = str_to_lower(group), #modify
  group = str_to_upper(group) #modify
)
```

## Arrange: in certain preferred order - ame as sorting

```{r}
arrange(litters_data, pups_born_alive)
arrange(litters_data, desc(pups_born_alive))
arrange(litters_data, pups_born_alive, gd0_weight)

```











